---
title: "Linear Functional Models"
subtitle: "Session 02"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=5, fig.height=3, fig.retina=3,
  out.width = "60%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Session's Goals

* Introduce the concept of mean change. <br/><br/>
* Introduce the linear functions. <br/><br/>
* Apply the mean change to characterize linear functions. <br/><br/>

---
# Motivation

> An online add publication service is offered by a company. At the physical store, the company offers the following prices: Publication during 7 days is of 186mxn  and for 14 days at a price of 237mxn. In the website, the company offers the following prices: Publication during 7 days is of 176.7mxn,  and for 14 days at a price of 226.1mxn.
--
<br/><br/>
1. If the service is not bought online, how much does the price change from 7 to 14 days? What if the service is bought online? <br/><br/>
1. Do you consider that both difference are "similar" or there is one better? <br/><br/>
1. How can we measure the difference between the prices for both services? <br/><br/>

---
# Mean Change

> __Definition:__ Given a dataset that contains the information (modeled by a function) we can measure the change of the function by the difference between the values of the function at some step: $$\frac{\mbox{change } y}{\mbox{change } x}=\frac{y_2-y_1}{x_2-x_1}=\frac{f(x_2)-f(x_1)}{x_2-x_1}$$<br/><br/>

---
# Linear functions

> __Definition:__ A function $f$ is said to be linear if it can be written as $f(x)=mx+b,$ where $m$ and $b$ are constants. <br/><br/>

--
<br/><br/>
1. $m$ is called the slope of the function.
2. $b$ is called the intercept of the function.

---
## How linear data looks like?

```{r, echo=FALSE, fig.cap="Graph of the function $f(x)=-3x+2$"}
library(ggplot2)
ggplot(data.frame(x=c(-5,5)), aes(x)) + 
  stat_function(fun = function(x) -3*x+2, geom = "line") +
  labs(x = "x", y = "f(x)")
```

---
#### Example 1

Given the function $f(x)=3x+2,$ compute value of $f$ on $x=-3,-1,0,1,3$ and then plot the function.

--
```{r}
x <- c(-3,-1,0,1,3)
y <- 3*x+2
plot(x,y)
```
---
#### Example 2

Given the function $f(x)=-4x+1,$ compute value of $f$ on $x=-2,-1,0,1,2$ and then plot the function.

--
```{r}
x <- c(-2,-1,0,1,2)
y <- -4*x+1
plot(x,y)
```

---
## What is the mean change of a linear function?

> __Definition:__ Given a linear function $f(x)=mx+b,$ the mean change of $f$ is $m.$ <br/><br/>

--
#### Example 3

Convince yourself that the mean change of the function $f(x)=3x+2$ is $3.$

---
#### How to determine if our data fits a linear function model?

```{r}	
tab <- as.table(rbind(c(327, 468,500,762), c(239, 477,382,472)))
dimnames(tab) <- list(data = c("x", "f(x)"),
                      index  = c("1", "2", "3","4"))
tab

```
---
#### How to determine if our data fits a linear function model?

> __Criterion:__ If the mean change of the function is constant, then the data fits a linear function model. <br/><br/>

--
<br/><br/>
In the previous data, we have that the mean changes between steps are:

$$mc_1=\frac{477-239}{468-327}=1.68$$
$$mc_2=\frac{382-477}{500-468}=-2.96$$
$$mc_3=\frac{472-382}{762-500}=0.34$$

---
#### Example 4

Determine if the following data fixed a linear model

```{r}	
tab <- as.table(rbind(c(3, 6,9,12), c(8,17,26,35)))
dimnames(tab) <- list(data = c("x", "f(x)"),
                      index  = c("1", "2", "3","4"))
tab

```
