---
title: "Mean and Instant Velocity"
subtitle: "Session 01"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

# Goals:

 * Understand the differences between mean and instant velocity. <br/><br/>
 * The derivative as a instant velocity indicator. <br/><br/>
 * The vectorial version of displacement and velocity.
---
# Problematic.

> A body falls from a building of 500 m height It is evident, body's height $h$ diminish by the time pass. 


> Note that even if $h$ diminish, the _rate of change_ it is not constant. This is true because the falling _velocity_ is not constant. Why?

---
# Problematic, a graphic representation.

---
# The mean velocity.


---
# Computing the mean velocity.

> These are some known values for the height at the time $t$ in seconds.


---
# Instant velocity.


---
# Mathematical formulation for Instant velocity.

> Mean velocity is $$\overline{v}=\frac{\Delta h}{\Delta t}=\frac{\mbox{displacement}}{\mbox{time}}$$

If we reduce the time passed to cero (in the limit), we obtain the instant velocity.

$$v=\limit_{\Delta t\to 0}\overline{v}=\limit_{\Delta t\to 0}\frac{\Delta h}{\Delta t}=\frac{d h}{dt}.$$

---
# The vectorial version.

> The __vectorial curve__ that describe the displacement of the body is written as: $$\vec{r}(t)=t\mathbf{i}+(500-4.9t^2)\mathbf{j}.$$ <br/>
In this curve, we think that the $x-$axis correspond to tiem and the $y-$axis to the height. <br/><br/> Note that $\mathbf{i}$ and $\mathbf{j}$ are the canonical vectors.

---
# And the derivative?


> Well, we can translate the velocity (derivative) of the body using vectors. The derivative will be the derivative of each component of the displacement vector respect to time. $$\vec{v}(t)=\frac{d t}{dt}\mathbf{i}+\frac{d(500-4.9t^2)}{dt}\mathbf{j}=\mathbf{i}+-9.8t\mathbf{j}.$$

---
