---
title: "Integration Techniques and Applications"
subtitle: "Session 03"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

# Goals:

 * Direct integration. <br/><br/>
 * Substitution. <br/><br/>

---
# Direct Integration

> __Method:__ Direct integration is a method of integration in which we use the antiderivative of the integrand to find the value of the definite integral. <br/><br/>

--
#### Example

1. $\int_{0}^{1} (x^2 + 2x + 1) dx$ <br/><br/>
2. $\int_{-1}^{1} (x^3 - 2x^2 + 3x - 1) dx$ <br/><br/>
3. $\int_{-5}^{6} e^{x}dx$ <br/><br/>
4. $\int_{1}^{e^2} \frac{1}{x} dx$ <br/><br/>
5. $\int_{0}^{\pi} \sin(x) dx$ <br/><br/>

---
# Substitution

> __Method:__ Substitution is a method of integration in which we use a change of variable to transform the integrand into a form that is easier to integrate (and use direct integration). <br/><br/>

--
#### Example

1. $\int_{0}^{1} \frac{2x+2}{(x^2 + 2x + 1)} dx$ <br/><br/>
2. $\int_{-1}^{1} xe^{3x^2} dx$ <br/><br/>
3. $\int_{0}^{\pi/4} \tan(x)dx$ <br/><br/>

---
# Applications

