---
title: "Non-homogeneus Linear ODEs"
subtitle: "Session 02"  
author: 
  - "Alejandro Ucan"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#000080",
  secondary_color = "#f2f3f4",
  inverse_header_color = "#00147e"
)
```

```{r xaringan-editable, echo=FALSE}
xaringanExtra::use_editable(expires = 1)
xaringanExtra::use_scribble()
```

```{r xaringan-logo, echo=FALSE}
xaringanExtra::use_logo(
  image_url = "https://github.com/alxcn/TecLogoEIC/blob/9562a53875418e749a296c85808a19c85fc4be74/IngenieriaCiencias_Horizontal_RGB.png",
  position = xaringanExtra::css_position(top = "2em", right = "2em")
)
```
# Session's Goals

* Learn the generic form of a linear ODE <br/><br/>
* Learn how to solve linear ODEs <br/><br/>
* Learn the method of variation of parameters <br/><br/>

---
# Linear ODE.

> The generic form of a linear ODE is $$a_1(x)y'+a_0(x)y+g(x)=0,$$ where $a_0(x)$ and $a_1(x)$ are the coefficients. <br/><br/>
If $g(x)=0,$ then the ODE is said to be __homogeneous__, otherwise it is __non-homogeneous.__

--
##### Example:

Homogeneuos: $y'-2xy-2=0,\, xy'-4y-x^6e^x=0.$ <br/><br/>
Non-homogeneuos: $y'-2xy-2=1,\, xy'-4y-x^6e^x=1.$
---

## How to solve those ODEs?

> Put the ODE in form $$y'+P(x)y=f(x).$$ Our solutions have to be in an interval $I$ where $P(x)$ and $f(x)$ are continuous functions. <br/><br/>
The solution is given by $$y(x)=y_h(x)+y_p(x),$$ where $y_h$ is a solution to the homogeneous version of the ODE and $y_p$ is a particular solution of the ODE.

---
## Method: Integrating Factor

> We need to find a solution of the form $y_p=u(x)y_1(x)$ where $y_1$ is a solution of the homogeneous version of the ODE. <br/><br/>
Método:
  1. Escribir la EDO de la forma: $$y'+P(x)y=f(x).$$
  1. The coefficient function $P(x)$ will help us to compute our __integrating factor__ $$e^{\int P(x)dx}.$$
  1. If we multiply our ODE by the integrating factor, we obtain $$\frac{d}{dx}\left[e^{\int P(x)dx} y\right]= e^{\int P(x)dx}f(x).$$
  1. Integrate in both sides and solve for $y.$
---
### Example 1

> Solve $4y'-3y=4.$

1. Putting the ODE in the form $y'+P(x)y=f(x)$ we have $y'-\frac{3}{4}y=1.$

1. $P(x)=-\frac{3}{4},$ therefore the integrating factor $$e^{\int \frac{-3}{4}dx}=e^{\frac{-3}{4}x}.$$

1. We multiply our ODE by the integrating factor and we obtain $$\frac{d}{dx}\left[e^{\frac{-3}{4}x}y\right]=e^{\frac{-3}{4}x}.$$

1. We integrate and solve for $y.$ $$e^{\frac{-3}{4}x}y=c\Rightarrow y=ce^{\frac{3}{4}x}.$$
Integramos y despejamos: $$e^{-3x}y=c\Rightarrow y=c e^{3x}.$$

---
### Example 2

> Solve $y'-2y=6.$

1. It is already in its standard form.

1. $P(x)=-2,$ therefore the integrating factor is $$e^{\int -2dx}=e^{-2x}.$$

1. We multiply our ODE by the integrating factor and we obtain $$\frac{d}{dx}\left[e^{-2x}y\right]=6e^{-2x}.$$

1. We integrate both sides and solve for $y$: $$e^{-2x}y=-3e^{-2x}+c\Rightarrow y=-3+ce^{-2x}.$$

---
### Ejemplo 3

> Resuelva la ecuación $y'-2xy=2x.$ 

---
# Non-free Falling Body

> Consider a falling body but the medium where it falls presents a resistance (let's say the air). If the velocity with which the body falls is given by $v(t)$ then it satisfies the non-homogeneous linear ODE: $$mv'+\beta v=mg,$$ where $m$ is the mass of the object, $\beta$ is the resistance constant, and $g$ is the gravity force.
---
## Lets solve it!

Using __integrating factor__ we can solve this ODE, let's see:

1. $P(x)=\beta/m,$ $f(x)=g$ y the integrating factor: $$e^{\frac{\beta t}{m}}.$$

1. Multiplying the integrating factor and transform the equation $$\frac{d}{dt}\left[e^{\frac{\beta t}{m}}v\right]=ge^{\frac{\beta t}{m}}.$$

1. Integrating both sides and solve for $v$: $$v(t)=\frac{mg}{\beta}+ce^{-\frac{\beta t}{m}}.$$

